\section{Theoretical Analysis}
\label{appendix:proof}
\subsection{Design Rationale}

The central design question is how the divergence signal should enter optimization.
We adopt the reverse-KL-aligned gap
\[
\Delta_t = \log \pi_T(y_t \mid s_t^{+}) - \log \pi_{\theta}(y_t \mid s_t)
\]
rather than forward KL, because it naturally evaluates on student-sampled tokens
and avoids the computationally expensive full-vocabulary matching.
However, using this raw gap directly as a coefficient would create overly strong,
unbounded token-level gradients during early training or under severe teacher-student mismatch.
To resolve this, we wrap the gap in a sigmoid function
\[
g_t = \sigma(\beta \Delta_t),
\]
which transforms the raw discrepancy into a bounded and monotone importance weight
\[
g_t \in (0,1), \qquad \frac{\partial g_t}{\partial \Delta_t} > 0.
\]
This preserves the ordering of token importance while strictly preventing gradient explosion.
Finally, we apply a stop-gradient operator to the gate.
Detaching $g_t$ ensures it acts purely as a confidence weight
rather than creating an additional, self-referential optimization pathway,
yielding a stable, first-order weighted likelihood update.

\subsection{Theoretical Properties}

We formalize the stability and curriculum properties of \methodname{} through the following propositions.

\begin{proposition}[Equivalent Weighted Likelihood Form]
\label{prop:weighted_mle}
Assume that both $\log \pi_T(y_t \mid s_t^{+})$ and $g_t$ are detached from gradient computation.
Minimizing $\mathcal{L}_{\methodname}$ is equivalent, up to an additive constant,
to maximizing a token-weighted log-likelihood objective on student-sampled tokens:
\[
\mathcal{L}_{\methodname}
= C - \operatorname{Agg}\!\left( g_t \log \pi_{\theta}(y_t \mid s_t) \right),
\]
where
\[
C = \operatorname{Agg}\!\left( g_t \log \pi_T(y_t \mid s_t^{+}) \right)
\]
is constant with respect to $\theta$.
\end{proposition}

\begin{proof}
By definition,
\[
\mathcal{L}_{\methodname}
= \operatorname{Agg}\!\left(
g_t \bigl(\log \pi_T(y_t \mid s_t^{+}) - \log \pi_{\theta}(y_t \mid s_t)\bigr)
\right)
\]
\[
= \operatorname{Agg}\!\left(
g_t \log \pi_T(y_t \mid s_t^{+})
\right)
-
\operatorname{Agg}\!\left(
g_t \log \pi_{\theta}(y_t \mid s_t)
\right)
\]
\[
= C - \operatorname{Agg}\!\left( g_t \log \pi_{\theta}(y_t \mid s_t) \right),
\]
where the first term $C = \operatorname{Agg}\!\left( g_t \log \pi_T(y_t \mid s_t^{+}) \right)$ is constant w.r.t.\ $\theta$ since both $g_t$ and $\log \pi_T(y_t \mid s_t^{+})$ are detached.
\end{proof}

\begin{proposition}[Gradient Form]
\label{prop:grad_form}
Under the same assumptions,
the gradient of $\mathcal{L}_{\methodname}$ is strictly modulated by the bounded scalar gate:
\[
\nabla_{\theta}\mathcal{L}_{\methodname}
= - \operatorname{Agg}\!\left( g_t \nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t) \right).
\]
\end{proposition}

\begin{proof}
From Proposition~\ref{prop:weighted_mle},
\[
\nabla_{\theta}\mathcal{L}_{\methodname}
= \nabla_{\theta}\left[
C - \operatorname{Agg}\!\left( g_t \log \pi_{\theta}(y_t \mid s_t) \right)
\right]
\]
\[
= 0 - \operatorname{Agg}\!\left( g_t \nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t) \right)
= - \operatorname{Agg}\!\left( g_t \nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t) \right).
\]
\end{proof}

\begin{proposition}[Monotonicity and Smoothness of the Gate]
\label{prop:gate}
The gate $g_t = \sigma(\beta \Delta_t)$ is strictly increasing in $\Delta_t$,
inducing an online token-level curriculum where larger discrepancies receive stronger weights.
Its derivative satisfies
\[
\frac{\partial g_t}{\partial \Delta_t}
= \beta \,\sigma(\beta \Delta_t)\bigl(1-\sigma(\beta \Delta_t)\bigr)
\in (0,\,\beta/4].
\]
\end{proposition}

\begin{proof}
By the chain rule,
\[
\frac{\partial g_t}{\partial \Delta_t}
= \beta \,\sigma'(\beta \Delta_t).
\]
Since the logistic sigmoid satisfies
\[
\sigma'(z) = \sigma(z)\bigl(1-\sigma(z)\bigr) > 0
\qquad \forall\, z \in \mathbb{R},
\]
we obtain
\[
\frac{\partial g_t}{\partial \Delta_t}
= \beta \,\sigma(\beta \Delta_t)\bigl(1-\sigma(\beta \Delta_t)\bigr)
> 0.
\]
Let $u = \sigma(\beta \Delta_t) \in (0,1)$:
\[
u(1-u) \le \left(\frac{u + (1-u)}{2}\right)^{\!2} = \frac{1}{4}
\]
\[
\frac{\partial g_t}{\partial \Delta_t}
= \beta\, u(1-u) \le \frac{\beta}{4}.
\]
\end{proof}

\begin{proposition}[Bounded Auxiliary Gradient]
\label{prop:bounded_grad}
Assume that $\|\nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t)\| \le B_t$ for each valid token.
Then the gate cannot amplify the auxiliary gradient beyond the unweighted likelihood gradient:
\[
\left\| \nabla_{\theta}\mathcal{L}_{\methodname} \right\|
\le \operatorname{Agg}(B_t).
\]
\end{proposition}

\begin{proof}
By Proposition~\ref{prop:grad_form},
\[
\left\| \nabla_{\theta}\mathcal{L}_{\methodname} \right\|
= \left\| \operatorname{Agg}\!\left( g_t \nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t) \right) \right\|
\]
\[
\le \operatorname{Agg}\!\left( g_t \left\| \nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t) \right\| \right)
\]
\[
\le \operatorname{Agg}\!\left( 1 \cdot B_t \right)
= \operatorname{Agg}(B_t),
\]
where the first inequality is the triangle inequality and the second uses $0 < g_t < 1$ and $\|\nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t)\| \le B_t$.
\end{proof}

\begin{proposition}[Effect of Not Detaching the Gate]
\label{prop:detach}
Without stop-gradient on the gate,
the non-detached token loss $\tilde{\ell}_t = \sigma(\beta \Delta_t)\,\Delta_t$
introduces an unstable self-referential coupling term into the gradient:
\[
\nabla_{\theta}\tilde{\ell}_t
= - \Bigl( g_t + \beta \Delta_t\, g_t(1-g_t) \Bigr)
\nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t).
\]
\end{proposition}

\begin{proof}
Write $\tilde{\ell}_t = g_t \Delta_t$. Since $\log \pi_T(y_t \mid s_t^{+})$ is constant w.r.t.\ $\theta$,
\[
\nabla_{\theta}\Delta_t
= \nabla_{\theta}\bigl[\log \pi_T(y_t \mid s_t^{+}) - \log \pi_{\theta}(y_t \mid s_t)\bigr]
= -\nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t).
\]
By the chain rule on $g_t = \sigma(\beta \Delta_t)$,
\[
\nabla_{\theta}g_t
= \beta\, \sigma'(\beta \Delta_t)\,\nabla_{\theta}\Delta_t
= \beta\, g_t(1-g_t)\,\nabla_{\theta}\Delta_t
= -\beta\, g_t(1-g_t)\,\nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t).
\]
Applying the product rule,
\[
\nabla_{\theta}\tilde{\ell}_t
= (\nabla_{\theta}g_t)\,\Delta_t + g_t\,(\nabla_{\theta}\Delta_t)
\]
\[
= \bigl[-\beta\, g_t(1-g_t)\,\nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t)\bigr]\,\Delta_t
  + g_t\,\bigl[-\nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t)\bigr]
\]
\[
= -\beta\, g_t(1-g_t)\,\Delta_t\,\nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t)
   - g_t\,\nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t)
\]
\[
= -\Bigl( g_t + \beta \Delta_t\, g_t(1-g_t) \Bigr)\,
   \nabla_{\theta}\log \pi_{\theta}(y_t \mid s_t).
\]
\end{proof}
